# Section 5.8.1 Greedy Matching and Subsequent Analysis of Hazard Rates (SD & PS)

```{r message=F, warning=F, error=F}
library(MatchIt)
library(cobalt)
library(gmodels)
library(ggplot2)
library(survival)
library(survMisc)
library(survminer)
library(knitr)
library(kableExtra)
library(tidyverse)
select <- dplyr::select
```

## Preparation

```{r}
# load data
d <- haven::read_dta("data/Data_Chapter5/chpt5_1_original.dta")

# make convenient objects
## names of match vars
match_vars1 <- colnames(select(d, PSH17A:other, -aodserv))
match_vars2 <- colnames(select(d, PSH17A:cwwrep, -aodserv))
match_vars3 <- colnames(select(d, PSH17A:cgneed, -aodserv))

## PS formula
psform1 <- f.build("aodserv", match_vars1)
psform2 <- f.build("aodserv", match_vars2)
psform3 <- f.build("aodserv", match_vars3)

# Function for balance check
get_bal <- function(out)
  cobalt::love.plot(out,
    binary = "std",
    stats = c("mean.diffs"),
    threshold = c(.1),
    var.order = "unadjusted",
    line = TRUE
)
```

## Propensity Score Estimation 

```{r}
# logistic 1
psmodel1 <- glm(psform1, data = d, family = binomial())
summary(psmodel1)
d$p1 <- predict(psmodel1, newdata = d, type = "response")
d$ps1 <- log((1 - d$p1) / (d$p1))

# logistic 2
psmodel2 <- glm(psform2, data = d, family = binomial())
summary(psmodel2)
d$p2 <- predict(psmodel2, newdata = d, type = "response")
d$ps2 <- log((1 - d$p2) / (d$p2))

# logistic 3
psmodel3 <- glm(psform3, data = d, family = binomial())
summary(psmodel3)
d$p3 <- predict(psmodel3, newdata = d, type = "response")
d$ps3 <- log((1 - d$p3) / (d$p3))
```

## Bivariate Comparisons

```{r warning=F, error=F, message=F}
d %>%
  select(married, educ, pov, employ, open, race, chdage, cgage, CRA47A, mental, 
         arrest, PSH17A, maltx, ra, cidi, cgneed, cwwrep, aodserv) %>%
  pivot_longer(-aodserv, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(bivariate.test = map(data, ~chisq.test(.$aodserv, .$value, correct = F))) %>%
  mutate(statistic = map(bivariate.test, ~round(.$statistic, 3))) %>%
  mutate(p.value = map(bivariate.test, ~round(.$p.value, 3))) %>%
  unnest(cols = c(statistic, p.value)) %>%
  select(variable, statistic, p.value)
```

Bivariate tests with contingency tables can also be accomplished with:

```{r eval=F}
gmodels::CrossTable(d$married, d$aodserv, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

## PSM based on Logistic regression & Covariate Balcance Check   

**Note: The matching results here are slightly different from the results in STATA. With R, we also prefer to use plots to check balances, which is different from the approach in STATA.**

### Scheme 1 - Nearest neighbor within caliper .25*SD

```{r message=F, warning=F, error=F}
# PSM based on Logistic regression 1 & Covariate Balance Check
set.seed(100)
m_nc1_1.out <-
  matchit(psform1,
    data = d,
    method = "nearest",
    distance = d$ps1,
    m.order = "random",
    caliper = .25,
    replace = T
  )
m_nc1_1.out
m_nc1_1.data <- MatchIt::match.data(m_nc1_1.out)
get_bal(m_nc1_1.out)
```

### Scheme 2 - Nearest neighbor within caliper .1*SD 

```{r message=F, warning=F, error=F}
# Note: Unlike Stata, the default caliper setting in matchit() is in standard deviation units (i.e., caliper = 0.1 is 0.1*SD)
set.seed(100)
m_nc1_2.out <-
  matchit(psform1,
    data = d,
    method = "nearest",
    distance = d$ps1,
    m.order = "random",
    caliper = .1,
    replace = T
  )
m_nc1_2.out
m_nc1_2.data <- match.data(m_nc1_2.out)
get_bal(m_nc1_2.out)
```

### Scheme 3 - Mahalanobis without propensity score

```{r message=F, warning=F, error=F}
set.seed(100)
m_m1.out <- matchit(psform1,
  data = d,
  distance = "mahalanobis",
  m.order = "random"
)
m_m1.out
m_m1.data <- match.data(m_m1.out)
get_bal(m_m1.out)
```

### Scheme 4 - Mahalanobis with propensity score

```{r message=F, warning=F, error=F}
match_vars1m <- colnames(select(d, PSH17A:other, ps1, -aodserv))
psform1m <- f.build("aodserv", match_vars1m)
set.seed(100)
m_mp1.out <- matchit(psform1m,
  data = d,
  distance = "mahalanobis"
)
m_mp1.out
m_mp1.data <- match.data(m_mp1.out)
get_bal(m_mp1.out)
```

## PSM based on Logistic regression 2

### Scheme 5 - Nearest neighbor within caliper .25*SD

```{r message=F, warning=F, error=F}
set.seed(100)
m_nc2_1.out <-
  matchit(psform2,
    data = d,
    method = "nearest",
    distance = d$ps2,
    m.order = "random",
    caliper = .25,
    replace = T
  )
m_nc2_1.out
m_nc2_1.data <- match.data(m_nc2_1.out) 
get_bal(m_nc2_1.out)
```

### Scheme 6 - Nearest neighbor within caliper .1*SD (Note: caliper in MatchIt is not real data)

```{r message=F, warning=F, error=F}
set.seed(100) 
m_nc2_2.out <-
  matchit(psform2, 
    data = d,
    method = "nearest", 
    distance = d$ps2, 
    m.order = "random", 
    caliper = .1,
    replace = T
  )
m_nc2_2.out
m_mp2.data <- match.data(m_nc2_2.out)
get_bal(m_nc2_2.out)
```

### Scheme 7 - Mahalanobis without propensity score

```{r message=F, warning=F, error=F}
set.seed(100)
m_m2.out <- matchit(psform2,
  data = d,
  distance = "mahalanobis",
  m.order = "random"
)
m_m2.out
m_m2.data <- match.data(m_m2.out) 
get_bal(m_m2.out)
```

### Scheme 8 - Mahalanobis with propensity score 

```{r eval=F}
vars2m <- colnames(select(d, PSH17A:cwwrep, ps2, -aodserv))
psform2m <- f.build("aodserv", vars2m)

set.seed(100)
m_mp2.out <- matchit(psform2m,
  data = d,
  distance = "mahalanobis"
)
m_mp2.out
m_mp2.data <- match.data(m_mp2.out)
get_bal(m_mp2.out)
```

## PSM based on Logistic regression 3 

### Scheme 9 - Nearest neighbor within caliper .25*SD

```{r message=F, warning=F, error=F}
set.seed(100) 
m_nc3_1.out <-
  matchit(psform3, 
    data = d,
    method = "nearest", 
    distance = d$ps3, 
    m.order = "random", 
    caliper = .25,
    replace = T
  )
m_nc3_1.out
m_nc3_1.data <- match.data(m_nc3_1.out) 
get_bal(m_nc3_1.out)
```

### Scheme 10 - Nearest neighbor within caliper .1*SD

```{r message=F, warning=F, error=F}
set.seed(100) 
m_nc3_2.out <-
  matchit(psform3, 
    data = d,
    method = "nearest", 
    distance = d$ps3, 
    m.order = "random", 
    caliper = .1,
    replace = T
  )
m_nc3_2.out
m_nc3_2.data <- match.data(m_nc3_2.out) 
get_bal(m_nc3_2.out)
```

### Scheme 11 - Mahalanobis without propensity score

```{r message=F, warning=F, error=F}
set.seed(100)
m_m3.out <- matchit(psform3,
  data = d,
  distance = "mahalanobis",
  m.order = "random"
)
m_m3.out
m_m3.data <- match.data(m_m3.out) 
get_bal(m_m3.out)
```

### Scheme 12 - Mahalanobis with propensity score

```{r message=F, warning=F, error=F}
match_vars3m <- colnames(select(d, PSH17A:cgneed, ps3, -aodserv))
psform3m <- f.build("aodserv", match_vars3m)
set.seed(100)
m_mp3.out <- matchit(psform3m,
  data = d,
  distance = "mahalanobis"
)
m_mp3.out
m_mp3.data <- match.data(m_mp3.out)
get_bal(m_mp3.out)
```

## Bivariate test after matching

```{r message=F, warning=F, error=F}
# Check balance using chi-squared tests
check_balance <- function(df) {

  # List of covariates to check balance
  covs <- c("married", "high", "bahigh",  "poverty2", "poverty3",
            "poverty4", "poverty5", "employ", "open", "black", "hispanic",
            "natam",  "cgrage1", "cgrage2", "cgrage3", 
            "chdage1", "chdage2", "chdage3", "CRA47A", "mental",  
            "arrest", "PSH17A", "sexual", "provide", "supervis", 
            "other",  "ra", "cidi", "cgneed", "cwwrep")
  
  # Function to calculate chisq.test p-values
  calc <- function(cov) {
    return(data.frame(covariate = cov,
                  p = chisq.test(df[cov], df$aodserv)$p.value))
  }
  
  # Calculate chisq.test for each covariate
  results <- map_dfr(covs, calc) %>%
    mutate(sig = ifelse(p < 0.05, "sig", "not sig")) %>%
    mutate(p = round(p, 2))
  
  # Get covariates that did not overlap sufficiently (p < .05)
  results_sig <- results %>%
    filter(sig == "sig")
  
  # Return results
  return(list(table = results, sig = toupper(results_sig$covariate)))
    
}
```

```{r message=F, warning=F, error=F}
# Combine tables 5.5 and 5.6
get_table <- function(scheme, df) {
  sig_vars <- ifelse(is.data.frame(df), paste(check_balance(df)$sig, collapse = ", "), df)
  treated_n <- ifelse(is.data.frame(df), nrow(filter(df, aodserv == 1)), NA)
  control_n <- ifelse(is.data.frame(df), nrow(filter(df, aodserv == 0)), NA)
  results <- tibble(Scheme = scheme,
                    SigCovariates = sig_vars,
                    Treated = treated_n,
                    Control = control_n)
}
table_schemes <- c("Original", "1. Nearest 1-1", "2. Nearest 1-2", 
                       "3. Mahalanobis 1", "4. Mahalanobis 1 with p-score added",
                       "5. Nearest 2-1", "6. Nearest 2-2",
                       "7. Mahalanobis 2", "8. Mahalanobis 2 with p-score added",
                       "9. Nearest 3-1", "10. Nearest 3-2",
                       "11. Mahalanobis 3", "12. Mahalanobis 3 with p-score added")
table_dfs <- list(d, m_nc1_1.data, m_nc1_2.data, # Original, Schemes 1, 2
                      m_m1.data, m_mp1.data, # Schemes 3, 4
                      m_nc2_1.data, m_mp2.data, # Schemes 5, 6
                      m_m2.data, m_mp2.data , # Schemes 7, 8
                      m_nc3_1.data, m_nc3_2.data, # Schemes 9, 10
                      m_m3.data, m_mp3.data) # Schemes 11, 12
table_summary <- map2_dfr(table_schemes, table_dfs, get_table)
options(knitr.kable.NA = "") # Do not display NAs in kable
table_summary %>%
  kbl(format = "html", 
      col.names = c("Scheme", 
 "Covariate Distributions Did Not Overlap Sufficiently: Covariates Significant After Matching p < .05",
 "Treated", "Control")) %>%
  add_header_above(c(" " = 2, "N of the New Sample" = 2))
```

## Kaplan-Meier and significance test of difference on survivor function

### All

```{r message=F, warning=F, error=F}
# all
ds <- haven::read_dta("data/Data_Chapter5/svl_all.dta")

## KM survival analysis
fit <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds)
fit
quantile(fit, probs = 0.15)

## Log-rank or Mantel-Haenszel test
fit.test <- survival::survdiff(Surv(durm, rrpt) ~ aodserv, data = ds, rho = 0)
fit.test
comp(ten(fit)) # the first n row is the Gehan-Breslow generalized Wilcoxon test

## plot 
ggsurvplot(fit, data = ds, pval = F)
```

### Scheme 1

```{r message=F, warning=F, error=F}
ds1 <- haven::read_dta("data/Data_Chapter5/svl_1.dta")

## KM survival analysis
fit1 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds1)
fit1
quantile(fit1, probs = 0.15)

## Log-rank or Mantel-Haenszel test
fit1.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds1, rho = 0)
fit1.test
comp(ten(fit1)) 

## plot
ggsurvplot(fit1, data = ds1, pval = F)
```

### Scheme 2

```{r message=F, warning=F, error=F}
ds2 <- haven::read_dta("data/Data_Chapter5/svl_2.dta")
## KM survival analysis
fit2 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds2)
fit2
quantile(fit2, probs = 0.15)

## Log-rank or Mantel-Haenszel test
fit2.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds2, rho = 0)
fit2.test
comp(ten(fit2)) 

## plot
ggsurvplot(fit2, data = ds2, pval = F)
```

### Scheme 3

```{r message=F, warning=F, error=F}
ds3 <- haven::read_dta("data/Data_Chapter5/svl_3.dta")
## KM survival analysis
fit3 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds3)
fit3
quantile(fit3, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit3.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds3, rho = 0)
fit3.test
comp(ten(fit3)) 

## plot
ggsurvplot(fit3, data = ds3, pval = F)
```

### Scheme 4

```{r message=F, warning=F, error=F}
ds4 <- haven::read_dta("data/Data_Chapter5/svl_4.dta")
## KM survival analysis
fit4 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds4)
fit4
quantile(fit4, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit4.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds4, rho = 0)
fit4.test
comp(ten(fit4)) 

## plot
ggsurvplot(fit4, data = ds4, pval = F)
```

### Scheme 5

```{r message=F, warning=F, error=F}
ds5 <- haven::read_dta("data/Data_Chapter5/svl_5.dta")
## KM survival analysis
fit5 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds5)
fit5
quantile(fit5, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit5.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds5, rho = 0)
fit5.test
comp(ten(fit5)) 

## plot
ggsurvplot(fit5, data = ds5, pval = F)
```

### Scheme 6

```{r message=F, warning=F, error=F}
ds6 <- haven::read_dta("data/Data_Chapter5/svl_6.dta")
## KM survival analysis
fit6 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds6)
fit6
quantile(fit6, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit6.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds6, rho = 0)
fit6.test
comp(ten(fit6)) 

## plot
ggsurvplot(fit6, data = ds6, pval = F)
```

### Scheme 7

```{r message=F, warning=F, error=F}
ds7 <- haven::read_dta("data/Data_Chapter5/svl_7.dta")
## KM survival analysis
fit7 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds7)
fit7
quantile(fit7, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit7.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds7, rho = 0)
fit7.test
comp(ten(fit7)) 

## plot
ggsurvplot(fit7, data = ds7, pval = F)
```

### Scheme 8

```{r message=F, warning=F, error=F}
ds8 <- haven::read_dta("data/Data_Chapter5/svl_8.dta")
## KM survival analysis
fit8 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds8)
fit8
quantile(fit8, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit8.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds8, rho = 0)
fit8.test
comp(ten(fit8)) 

## plot
ggsurvplot(fit8, data = ds8, pval = F)
```


### Scheme 9

```{r message=F, warning=F, error=F}
ds9 <- haven::read_dta("data/Data_Chapter5/svl_9.dta")
## KM survival analysis
fit9 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds9)
fit9
quantile(fit9, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit9.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds9, rho = 0)
fit9.test
comp(ten(fit9)) 

## plot
ggsurvplot(fit9, data = ds9, pval = F)
```

### Scheme 10

```{r message=F, warning=F, error=F}
ds10 <- haven::read_dta("data/Data_Chapter5/svl_10.dta")
## KM survival analysis
fit10 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds10)
fit10
quantile(fit10, probs = 0.15)
## Log-rank or Mantel-Haenszel test
fit10.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds10, rho = 0)
fit10.test
comp(ten(fit10)) 

## plot
ggsurvplot(fit10, data = ds10, pval = F)
```

### Scheme 11

```{r message=F, warning=F, error=F}
ds11 <- haven::read_dta("data/Data_Chapter5/svl_11.dta")
## KM survival analysis
fit11 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds11)
fit11
quantile(fit11, probs = 0.1)
## Log-rank or Mantel-Haenszel test
fit11.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds11, rho = 0)
fit11.test
comp(ten(fit11)) 

## plot
ggsurvplot(fit11, data = ds11, pval = F)
```

### Scheme 12

```{r message=F, warning=F, error=F}
ds12 <- haven::read_dta("data/Data_Chapter5/svl_12.dta")
## KM survival analysis
fit12 <- survfit(Surv(durm, rrpt) ~ aodserv, data = ds12)
fit12
quantile(fit12, probs = 0.1)
## Log-rank or Mantel-Haenszel test
fit12.test <- survdiff(Surv(durm, rrpt) ~ aodserv, data = ds12, rho = 0)
fit12.test
comp(ten(fit12)) 

## plot
ggsurvplot(fit12, data = ds12, pval = F)
```
