# Section 5.8.2, 5.8.3, and 5.8.4 (SD & PS)

## Section 5.8.2 Optimal Matching

```{r message=F, warning=F, error=F}
library(optmatch)
library(knitr)
library(broom)
library(sandwich)
library(kableExtra)
library(tidyverse)
library(rlang)
select <- dplyr::select
```

### Load Data

```{r message=FALSE, warning=FALSE}
d <- haven::read_dta("data/Data_Chapter5/chpt5_2_original.dta")
df <- haven::read_dta("data/Data_Chapter5/chpt5_2.dta")
cds <- haven::read_dta("data/Data_Chapter5/chpt5_2ps.dta")
```

### Bivariate Comparisons 

```{r message=FALSE, warning=FALSE}
# Intergenerational dependence 
gmodels::CrossTable(d$kuse, d$puse, 
                    prop.c = FALSE, prop.r = FALSE, prop.chisq = FALSE, 
                    chisq = TRUE)

# Wilcoxon Rank-Sum (Mann-Whitney) test and t-test
d %>%
  select(mratio96, pcged97, pcg_adc, black, age97, male, kuse) %>%
  pivot_longer(-kuse, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(wilcoxon = map(data, ~wilcox.test(.$value ~.$kuse, correct = F))) %>%
  mutate(wilcoxon.stat = map(wilcoxon, ~round(.$statistic, 3))) %>%
  mutate(wilcoxon.pvalue = map(wilcoxon, ~round(.$p.value, 3))) %>%
  unnest(cols = c(wilcoxon.stat, wilcoxon.pvalue)) %>%
  mutate(ttest = map(data, ~t.test(.$value ~ .$kuse, var.equal = T))) %>%
  mutate(t.stat = map(ttest, ~round(.$statistic, 3))) %>%
  mutate(t.pvalue = map(ttest, ~round(.$p.value, 3))) %>%
  unnest(cols = c(t.stat, t.pvalue)) %>%
  select(-data, -wilcoxon, -ttest)
```

### Outcome Without Bias Control

```{r message=FALSE, warning=FALSE}
# visual
d <- d %>% 
  dplyr::mutate(treatfac = case_when(kuse==0 ~ "Nonusers", 
                                      kuse==1 ~ "Users"))               
ggplot(d, aes(y = lwss97, x = treatfac)) + 
  geom_boxplot()

# t-test
t.test(lwss97 ~ kuse, data = d, var.equal = T) %>% 
  broom::tidy()

# regression 
reg <- lm( lwss97 ~ kuse + male + black + age97 + pcged97 + mratio96 + pcg_adc, 
           data = d)
broom::tidy(lmtest::coeftest(reg, vcov = vcovCL, cluster = d$pcg_id))
```

### Matching

*Note: To be consistent with the STATA result, we will use the propensity scores created by Stata.*

#### Distribution of estimated propensity scores

```{r message=FALSE, warning=FALSE}
cds <- cds %>% 
  dplyr::mutate( treatfac = case_when( kuse==0 ~ "Nonusers", 
                                       kuse==1 ~ "Users"))

ggplot(cds, aes(x = treatfac, y = ps)) + geom_boxplot() + 
  ggtitle("Box Plots of Estimated Propensity Scores") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(cds, aes(x = ps, color = treatfac)) + 
  geom_histogram(aes(y=..density..), fill="white", position="dodge")+
  geom_density(alpha=.2, fill="#FF6666") +
  ggtitle("Histograms of Estimated Propensity Scores") +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Optimal Matching

```{r message=FALSE, warning=FALSE, results="hide"}
attach(cds)
prank <- rank(ps)
names(prank) <- kid
d1 <- outer(prank[kuse==1], prank[kuse==0], "-")
d1 <- abs(d1)

#fullmatch fm
fm <- fullmatch(d1)
(fm.d <- matched.distances(fm,d1,pres=TRUE))
unlist(fm.d,max)

#variable match, (vm1 - at least 1 at most 4)
vm1 <- fullmatch(d1,min.controls=1,max.controls=4)
(vm1.d <- matched.distances(vm1,d1,pres=TRUE))
unlist(vm1.d,max)

#variable match, (vm2 - at least 2 at most 4)
vm2 <- fullmatch(d1,min.controls=2,max.controls=4)
(vm2.d <- matched.distances(vm2,d1,pres=TRUE))
unlist(vm2.d,max)

#variable match, (vm3 - use Hensen's equaion)
vm3 <- fullmatch(d1,min.controls=1.33,max.controls=5.32)
(vm3.d <- matched.distances(vm3,d1,pres=TRUE))
unlist(vm3.d,max)

#variable match, (vm4 at least 2 at most 7)
vm4 <- fullmatch(d1,min.controls=2,max.controls=7)
(vm4.d <- matched.distances(vm4,d1,pres=TRUE))
unlist(vm4.d,max)

#pairmatch pm
pm <- pairmatch(d1, controls=1)
(pm.d <- matched.distances(pm,d1,pres=TRUE))
unlist(pm.d,max)
```

```{r message=FALSE, warning=FALSE}
mean(unlist(fm.d))
sum(unlist(fm.d))
stratumStructure(fm)

mean(unlist(vm1.d))
sum(unlist(vm1.d))
stratumStructure(vm1)

mean(unlist(vm2.d))
sum(unlist(vm2.d))
stratumStructure(vm2)

mean(unlist(vm3.d))
sum(unlist(vm3.d))
stratumStructure(vm3)

mean(unlist(vm4.d))
sum(unlist(vm4.d))
stratumStructure(vm4)

mean(unlist(pm.d))
sum(unlist(pm.d))
stratumStructure(pm)
```

#### Table 5.10. Covariate Imbalance Before and After Matching by Matching Scheme
```{r message=FALSE, warning=FALSE}
# Imbalance function
imbalance <- function(cov, method) {
  
  # get quosure
  cov_name <- cov
  method_name <- method
  cov <- rlang::parse_expr(cov)
  method <- rlang::parse_expr(method)
  
  # dx
  df2 <- df %>%
    group_by(kuse) %>%
    summarise(m_x := mean(!!cov), 
              sd_x := sd(!!cov), .groups = "drop")
  mxt = df2[2,2]
  mxc = df2[1,2]
  s2xt = df2[2,3]^2
  s2xc = df2[1,3]^2
  sx = sqrt((s2xt+s2xc)/2)
  dx = as.numeric(abs(mxt-mxc)/sx)
  
  # dxm
  if (method != "before") {
    df3 <- df %>%
      group_by(!!method, kuse) %>%
      summarise(m_x = mean(!!cov),
                sd_x = sd(!!cov),
                n = n(),
                .groups = "drop")
    
    mxc = as.numeric(mean(filter(df3, kuse == 0)$m_x))
    mxt = as.numeric(mean(filter(df3, kuse == 1)$m_x))
    dxm_num = abs(mxt-mxc)
    dxm = as.numeric(dxm_num/sx)
  
    return(tibble(cov = cov_name, method = method_name, dx = NA, dxm = dxm))
  
  } else {
    return(tibble(cov = cov_name, method = method_name, dx = dx, dxm = NA))
  }
}

# Define covariates
cov_labels <- tibble(cov = c("mratio96", "pcged97", "pcg_adc", 
                                      "black", "age97", "male"),
                         new = c("Ratio of family income to poverty line in 1996", 
                                 "Caregiver's education in 1997 (years of schooling)", 
                                 "Caregiver's number of years using AFDC in childhood", 
                                 "Child's race: African American (reference: other)", 
                                 "Child's age in 1997", "Child's gender: male (reference: female)"),
                         order = c(1, 2, 3, 4, 5, 6))

# Calculate dx and dxm using imbalance()
arg1 <- rep(c("mratio96", "pcged97", "pcg_adc", "black", "age97", "male"), each = 7)
arg2 <- rep(c("before", "fm", "vm1", "vm2", "vm3", "vm4", "pm"), length(arg1)/7)
table_5.10 <- map2_dfr(arg1, arg2, imbalance) %>%
  mutate(method = recode(method,
                         `before` = "Before Matching",
                          `fm` = "Full Matching", 
                         `vm1` = "Variable Matching 1 (at least 1, at most 4)", 
                         `vm2` = "Variable Matching 2 (at least 2, at most 4)",
                         `vm3` = "Variable Matching 3 (Hansen's equation)", 
                         `vm4` = "Variable Matching 4 (at least 2, at most 7)",
                         `pm` = "Pair matching")) %>%
  left_join(cov_labels, by = "cov") %>%
  arrange(order) %>%
  select(new, method, dx, dxm)

# Print Table 5.10
options(knitr.kable.NA = "")
knitr::kable(table_5.10, 
             col.names = c("Covariate", "Matching Scheme", "$d_x$", "$d_{xm}$"), 
             escape = FALSE,
             format = "html", 
             digits = 2,
             caption = "Table 5.10. Covariate Imbalance Before and After Matching by Matching Scheme") %>%
  kable_styling(latex_options = "striped") %>%
  kable_styling(position = "center") %>%
  collapse_rows(columns = 1, valign = "top")
```

## Section 5.8.3 Post-Full-Matching Analysis of Outcome

### Hodges-Lehmann Aligned Rank Test 

```{r message=FALSE, warning=FALSE}
# Shenyang Guo's Stata program, hodgesl, version 8.2
hodgesl <- function(dataname, varname, blockname, treatname) {
  blockname_str <- deparse(substitute(blockname))
  set.seed(1000)
  renamed_file <- dataname %>%
    filter(!is.na({{blockname}}))
  r1 <- renamed_file %>%
    group_by({{blockname}}) %>%
    summarise(m_y = mean({{varname}}), .groups = "drop") %>%
    arrange({{blockname}})
  r2 <- renamed_file %>%
    group_by({{blockname}}, {{treatname}}) %>%
    summarise(mean_y = mean({{varname}}),
              n = n(), .groups = "drop") %>%
    mutate(mean_diff = ifelse({{treatname}} == 1,
                              ((n+lag(n))/sum(n))*(mean_y-lag(mean_y)), NA)) %>%
    mutate(tx_effect = sum(mean_diff, na.rm = T)) %>%
    mutate(i = row_number()) %>%
    slice(1) %>%
    select(tx_effect, i)
  fm_results <- renamed_file %>%
    group_by({{blockname}}, {{treatname}}) %>%
    summarise(mean_y = mean({{varname}}),
              n = n(), .groups = "drop")
  r3 <- renamed_file %>%
    group_by({{blockname}}, {{treatname}}) %>%
    summarise(m_or_n = n(), .groups = "drop") %>%
    arrange({{blockname}}, {{treatname}}) %>%
    mutate(mi = ifelse({{treatname}} == 0, m_or_n, 
                ifelse({{treatname}} == 1, NA, NA))) %>%
    mutate(ni = ifelse({{treatname}} == 1, m_or_n,
                ifelse({{treatname}} == 0, NA, NA))) %>%
    mutate(Ni = ni+lag(mi)) %>%
    mutate(mi = ifelse(is.na(mi), lag(mi), mi)) %>%
    filter(!is.na(Ni)) %>%
    mutate(factor = (mi*ni)/(Ni*(Ni-1))) %>%
    select({{blockname}}, factor) %>%
    arrange({{blockname}})
  r4 <- renamed_file %>%
    arrange({{blockname}}) %>%
    left_join(r1, by = blockname_str) %>%
    mutate(dy = {{varname}}-m_y) %>%
    arrange(dy) %>%
    mutate(rk = row_number()) %>%
    arrange({{blockname}})
  r4a <- r4 %>%
    filter({{treatname}} != 0) %>%
    group_by({{blockname}}) %>%
    summarise(wsi = sum(rk), .groups = "drop")
  r5 <- r4 %>%
    group_by({{blockname}}) %>%
    summarise(ki_ = mean(rk), .groups = "drop")
  r6 <- r4 %>%
    filter({{treatname}} != 0) %>%
    group_by({{blockname}}) %>%
    summarise(ni = n(), .groups = "drop") %>%
    arrange({{blockname}}) %>%
    left_join(r5, by = blockname_str) %>%
    mutate(E_wsi = ni*ki_) %>%
    arrange({{blockname}})
  r7 <- r4 %>%
    arrange({{blockname}}) %>%
    left_join(r5, by = blockname_str) %>%
    mutate(k = (rk-ki_)^2) %>%
    group_by({{blockname}}) %>%
    summarise(ss_kd_i = sum(k), .groups = "drop") %>%
    arrange({{blockname}})
  results <- r3 %>%
    arrange({{blockname}}) %>%
    left_join(r7, by = blockname_str) %>%
    left_join(r6, by = blockname_str) %>%
    left_join(r4a, by = blockname_str) %>%
    mutate(var_wsi = factor*ss_kd_i,
          var = sum(var_wsi),
          sum_Ewsi = sum(E_wsi),
          ws = sum(wsi),
          HL_mean = ws-sum_Ewsi,
          HL_se = sqrt(var),
          z = HL_mean/HL_se,
          p = 1 - pnorm(abs(z))) %>%
    select(HL_mean, HL_se, z, p) %>%
    slice(1) %>%
    mutate(i = row_number()) %>%
    left_join(r2, by = "i")
  return(results)
}

df.h <- haven::read_dta("data/Data_Chapter5/chpt5_2(original).dta") %>%
  mutate(fm_s = paste0(round(fm, 5), "s"))
hodgesl(df.h, lwss97, fm_s, kuse)
```

### Cohen's d

```{r message=FALSE, warning=FALSE}
# Shenyang Guo's Stata program, imbalance, version 8.2
imbalance2 <- function(df, varname, treatname, blockname) {
  
  # dx
  df2 <- df %>%
    group_by({{treatname}}) %>%
    summarise(m_x = mean({{varname}}), sd_x = sd({{varname}}), .groups = "drop")
  mxt = df2[2,2]
  mxc = df2[1,2]
  s2xt = df2[2,3]^2
  s2xc = df2[1,3]^2
  sx = sqrt((s2xt+s2xc)/2)
  dx = as.numeric(abs(mxt-mxc)/sx)
  
  # dx
  df3 <- df %>%
    group_by({{blockname}}, {{treatname}}) %>%
    summarise(m_x = mean({{varname}}),
              sd_x = sd({{varname}}),
              n = n(),
              .groups = "drop")
    
  mxc = as.numeric(mean(filter(df3, {{treatname}} == 0)$m_x))
  mxt = as.numeric(mean(filter(df3, {{treatname}} == 1)$m_x))
  dxm_num = abs(mxt-mxc)
  dxm = as.numeric(dxm_num/sx)
  
  return(list(dx = dx, dxm = dxm))
}

df.h2 <- haven::read_dta("data/Data_Chapter5/chpt5_2(original).dta")
imbalance2(df.h2, lwss97, kuse, fm)$dxm # dxm for the outcome variable is Cohen's d
```

## Section 5.8.4 Post-Matching Analysis Using Regression of Difference Scores

```{r message=FALSE, warning=FALSE}
df <- haven::read_dta("data/Data_Chapter5/chpt5_2.dta")

# kuse == 1
df1 <- df %>% 
  select(pm, kuse, y1 = lwss97, male1 = male, black1 = black, age971 = age97, 
         pcged971 = pcged97, mratio961 = mratio96, pcg_id) %>% 
  filter(kuse==1) 

# kuse == 0
df0 <- df %>% 
  select(pm, kuse, y0 = lwss97, male0 = male, black0 = black, age970 = age97, pcged970 = pcged97, mratio960 = mratio96, pcg_id) %>% 
  filter( kuse == 0 ) 

# merge data
df2 <- left_join(df0, df1, "pm")
df2 <- df2 %>% 
  mutate(y = y1 - y0,
         male = male1 - male0,
         black = black1 - black0,
         age97 = age971 - age970,
         pcged97 = pcged971 - pcged970,
         mratio96 = mratio961 - mratio960)

# regression 
reg1 <- lm(y ~ male + black + age97 + pcged97 + mratio96, data = df2)
broom::tidy(lmtest::coeftest(reg1, vcov = vcovCL, cluster = df2$pcg_id.x))
```
