# Section 5.8.5 Multilevel Propensity Score (PS)

```{r message=F, warning=F, error=F}
library(tidyverse)
library(haven)
library(broom)
library(MatchIt)
library(cobalt)
library(plm)
library(lme4)
library(lmerTest)
library(sjlabelled)
library(performance)
options(scipen=999)
```

## cccpros57

### Import Data

```{r}
# Remove formats and labels, because matchit requires a vector for the distance option
df <- haven::read_dta("data/Data_Chapter5/cccpros57.dta") %>%
  zap_formats()
df <- sjlabelled::remove_all_labels(df) %>%
  as_tibble()

post1 <- haven::read_dta("data/Data_Chapter5/post1.dta")
post2 <- haven::read_dta("data/Data_Chapter5/post2.dta")
```

### Obtain Single-Level p-score

```{r}
m1 <- glm("intschb ~ agey + female + black + hisp + pcedu + incpovlr + pcempf + 
          fatherr + AYP05Cs + pmin05s", family = binomial, data = df)
m1.sum <- summary(m1)
exp(m1.sum$coefficients[,1])
exp(confint(m1))
ps_sgl <- predict(m1, type = "response") # equivalent to df$ps_sgl
```

### Obtain School-Level p-score Model

```{r}
# Keep the first row/observation for each schbl value
df.school <- df %>%
  arrange(schbl) %>%
  group_by(schbl) %>%
  slice(1)

m2 <- glm("intschb ~ AYP05Cs + pmin05s", family = binomial, data = df.school)
m2.sum <- summary(m2)
exp(m2.sum$coefficients[,1])
exp(confint(m2))
```

### Before-Matching Balance Check

```{r}
t.test(agey ~ intschb, data = df, var.equal = TRUE)
chisq.test(df$female, df$intschb, correct = FALSE)
chisq.test(df$black, df$intschb, correct = FALSE)
chisq.test(df$hisp, df$intschb, correct = FALSE)
t.test(pcedu ~ intschb, data = df, var.equal = TRUE)
t.test(incpovlr ~ intschb, data = df, var.equal = TRUE)
chisq.test(df$pcempf, df$intschb, correct = FALSE)
chisq.test(df$fatherr, df$intschb, correct = FALSE)
t.test(AYP05Cs ~ intschb, data = df, var.equal = TRUE)
t.test(pmin05s ~ intschb, data = df, var.equal = TRUE)

# Alternative balance check using matchit
scheme1.before <- MatchIt::matchit(intschb ~ agey + female + black + hisp + 
                    pcedu + incpovlr + pcempf + fatherr + AYP05Cs + pmin05s,
                    data = df,
                    method = NULL)
summary(scheme1.before)
```

### Matching with ps_sgl

Note that the caliper argument in `matchit()` is "the number of standard deviations of the distance measure within which to draw control units (default = 0, no caliper matching). If a caliper is specified, a control unit within the caliper for a treated unit is randomly selected as the match for that treated unit."^[https://r.iq.harvard.edu/docs/matchit/2.4-14/Additional_Arguments_f3.html]

```{r}
# Generate a column of x random numbers, sort x in ascending order, calculate caliper with ps_sgl
set.seed(1000)
df.scheme1 <- df %>%
  add_column(x = runif(nrow(.))) %>%
  arrange(x)

# Scheme 1 - Nearest Neighbor with ps_sgl
set.seed(1000)
scheme1 <- MatchIt::matchit(intschb ~ agey + female + black + hisp + pcedu + incpovlr +
                     pcempf + fatherr + AYP05Cs + pmin05s,
                   data = df.scheme1,
                   method = "nearest",
                   distance = df.scheme1$ps_sgl,
                   caliper = .25, # equivalent to sd(df.scheme1$ps_sgl)*.25
                   std.caliper = T, # caliper in standard deviation units
                   m.order = "largest", # descending order of distance measures
                   replace = FALSE)
sd(df.scheme1$ps_sgl)*.25 == scheme1$caliper # check caliper
scheme1.data <- MatchIt::match.data(scheme1)

# Balance Check
cobalt::love.plot(scheme1,
  binary = "std",
  stats = c("mean.diffs"),
  threshold = c(.1),
  var.order = "unadjusted",
  line = TRUE,
  s.d.denom = "pooled"
)


```

### Matching with ps_sch

```{r}
# Generate a column of x random numbers, sort x in ascending order, calculate caliper with ps_sch
set.seed(1000)
df.scheme2 <- df %>%
  add_column(x = runif(nrow(.))) %>%
  arrange(x)

# Scheme 2 - Nearest Neighbor with ps_sch
set.seed(1000)
scheme2 <- MatchIt::matchit(intschb ~ agey + female + black + hisp + pcedu + incpovlr +
                     pcempf + fatherr + AYP05Cs + pmin05s,
                   data = df.scheme2,
                   method = "nearest",
                   distance = df.scheme2$ps_sch,
                   caliper = .25, # equivalent to sd(df.scheme2$ps_sch)*.25
                   m.order = "largest", # descending order of distance measures
                   replace = FALSE)
scheme2.data <- MatchIt::match.data(scheme2)

# Balance Check
cobalt::love.plot(scheme2,
  binary = "std",
  stats = c("mean.diffs"),
  threshold = c(.1),
  var.order = "unadjusted",
  line = TRUE,
  s.d.denom = "pooled"
)
```

### Check ICC

```{r}
# Function to calculate Stata's "rho" (ICC) in xtreg
get_icc <- function(lmer_obj, group) {
  sigma_u2 <- as.numeric(attr(VarCorr(lmer_obj)[[group]], "stddev"))^2
  sigma_e2 <- summary(lmer_obj)$sigma^2
  return(sigma_u2/(sigma_u2+sigma_e2))
}

# Original dataset
m3 <- lmer(cccpros57 ~ (1|schbl), data = df, REML = F)
get_icc(m3, "schbl")

# Scheme 1: ICC using school as a clustering variable
m4 <- lmer(cccpros57 ~ (1|schbl), data = post1, REML = F)
get_icc(m4, "schbl")

# Scheme 1: ICC using the matched set as a clustering variable (create mset first)
m5 <- lmer(cccpros57 ~ (1|mset), data = post1, REML = F)
get_icc(m5, "mset")

# Scheme 2: ICC using school as a clustering variable
m6 <- lmer(cccpros57 ~ (1|schbl), data = post2, REML = F)
get_icc(m6, "schbl")

# Scheme 2: ICC using the matched set as a clustering variable (create mset first)
m7 <- lmer(cccpros57 ~ (1|mset), data = post2, REML = F)
get_icc(m7, "mset")
```

### Outcome Analsis: Multilevel

```{r}
out1 <- lmer(cccpros57 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = df,
           REML = FALSE)
summary(out1)

out2 <- lmer(cccpros57 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = post1,
           REML = FALSE)
summary(out2)

out3 <- lmer(cccpros57 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = post2,
           REML = FALSE)
summary(out3)

# Note that lmer() does not use the EM algorithm in Stata's xtmixed, but the results are similar
out4 <- lmer(cccpros57 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl) + (1|mset),
           data = post1,
           REML = FALSE,
           control=lmerControl(optCtrl=list(maxfun=30000)))
summary(out4)
```

## cccragg24

### Import Data

```{r}
ra <- haven::read_dta("data/Data_Chapter5/cccragg24.dta") %>%
  zap_formats()
ra <- sjlabelled::remove_all_labels(ra) %>%
  as_tibble()
post3 <- haven::read_dta("data/Data_Chapter5/post3.dta")
post4 <- haven::read_dta("data/Data_Chapter5/post4.dta")
```

### Obtain Single-Level p-score

```{r}
ra_glm <- glm("intschb ~ agey + female + black + hisp + pcedu + incpovlr + pcempf + 
          fatherr + AYP05Cs + pmin05s", family = binomial, data = ra)
ra_glm_sum <- summary(ra_glm)
exp(ra_glm_sum$coefficients[,1])
ra_ps_sgl <- predict(ra_glm, type = "response")
```

### Obtain School-Level p-score Model

```{r}
# Keep the first row/observation for each schbl value
ra_school <- ra %>%
  arrange(schbl) %>%
  group_by(schbl) %>%
  slice(1)

ra_glm2 <- glm("intschb ~ AYP05Cs + pmin05s", family = binomial, data = ra_school)
ra_glm2_sum <- summary(ra_glm2)
exp(ra_glm_sum$coefficients[,1])
```

### Before-Matching Balance Check

```{r}
t.test(agey ~ intschb, data = ra, var.equal = TRUE)
chisq.test(ra$female, ra$intschb, correct = FALSE)
chisq.test(ra$black, ra$intschb, correct = FALSE)
chisq.test(ra$hisp, ra$intschb, correct = FALSE)
t.test(pcedu ~ intschb, data = ra, var.equal = TRUE)
t.test(incpovlr ~ intschb, data = ra, var.equal = TRUE)
chisq.test(ra$pcempf, ra$intschb, correct = FALSE)
chisq.test(ra$fatherr, ra$intschb, correct = FALSE)
t.test(AYP05Cs ~ intschb, data = ra, var.equal = TRUE)
t.test(pmin05s ~ intschb, data = ra, var.equal = TRUE)
```

### Matching with ps_sgl

```{r}
# Scheme 1 - Nearest Neighbor with ps_sgl
set.seed(1000)
ra_scheme1 <- MatchIt::matchit(intschb ~ agey + female + black + hisp + pcedu + 
                                 incpovlr +pcempf + fatherr + AYP05Cs + pmin05s,
                               data = ra,
                               method = "nearest",
                               distance = ra$ps_sgl,
                               caliper = .25,
                               replace = FALSE)
ra_scheme1_data <- MatchIt::match.data(ra_scheme1)

# Balance Check
cobalt::love.plot(ra_scheme1,
  binary = "std",
  stats = c("mean.diffs"),
  threshold = c(.1),
  var.order = "unadjusted",
  line = TRUE,
  s.d.denom = "pooled"
)
```

### Matching with ps_sch

```{r}
# Scheme 2 - Nearest Neighbor with ps_sch
set.seed(1000)
ra_scheme2 <- MatchIt::matchit(intschb ~ agey + female + black + hisp + pcedu + incpovlr +
                pcempf + fatherr + AYP05Cs + pmin05s,
                data = ra,
                method = "nearest",
                distance = ra$ps_sch,
                caliper = .25,
                replace = FALSE)
ra_scheme2_data <- MatchIt::match.data(ra_scheme2)

# Balance Check
cobalt::love.plot(ra_scheme2,
  binary = "std",
  stats = c("mean.diffs"),
  threshold = c(.1),
  var.order = "unadjusted",
  line = TRUE,
  s.d.denom = "pooled"
)
```

### Check ICC

```{r}
i1 <- lmer(cccragg24 ~ (1|schbl), data = ra, REML = F)
get_icc(i1, "schbl")

i2 <- lmer(cccragg24 ~ (1|schbl), data = post3, REML = F)
get_icc(i2, "schbl")

i3 <- lmer(cccragg24 ~ (1|mset), data = post3, REML = F)
get_icc(i3, "mset")

i4 <- lmer(cccragg24 ~ (1|schbl), data = post4, REML = F)
get_icc(i4, "schbl")

i5 <- lmer(cccragg24 ~ (1|mset), data = post4, REML = F)
get_icc(i5, "mset")
```

### Outcome Analsis: Multilevel

```{r}
ra_out1 <- lmer(cccragg24 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = ra,
           REML = FALSE)
summary(ra_out1)

ra_out2 <- lmer(cccragg24 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = post3,
           REML = FALSE)
summary(ra_out2)

ra_out3 <- lmer(cccragg24 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl),
           data = post4,
           REML = FALSE)
summary(ra_out3)

# Note that lmer() does not use the EM algorithm in Stata's xtmixed
ra_out4 <- lmer(cccragg24 ~ intschb + agey + female + black + hisp + pcedu +
             incpovlr + pcempf + fatherr + AYP05Cs + pmin05s + (1|schbl) + (1|mset),
           data = post4,
           REML = FALSE,
           control=lmerControl(optCtrl=list(maxfun=30000)))
summary(ra_out4)
```
