# Section 5.8.6 Comparison of Rand-gbm and Stata-boost (PS) 

```{r message=F, warning=F, error=F}
library(tidyverse)
library(haven)
library(cobalt)
library(sjlabelled)
library(estimatr)
library(lmtest)
library(sandwich)
library(optmatch)
options(scipen=999)
select <- dplyr::select
```

## Import Data

```{r}
mc <- read_dta("data/Data_Chapter5/g3aca1_.dta") %>%
  zap_formats()
mc <- sjlabelled::remove_all_labels(mc) %>%
  as_tibble()
```

## Optimal pair matching Using pscores generated by Rand-gbm (ps)

```{r error=F, warning=F, message=F}
prank <- rank(mc$ps)
names(prank) <- mc$id
d1 <- outer(prank[mc$intbl==1], prank[mc$intbl==0], "-")
d1 <- abs(d1)
pm <- pairmatch(d1, controls=1)
(pm.d <- matched.distances(pm,d1,pres=TRUE))
unlist(pm.d,max)
mean(unlist(pm.d))
sum(unlist(pm.d))
stratumStructure(pm)

# Merge to original dataset and calculate difference scores
m1.df <- tibble(id = as.numeric(names(pm)), gbm_pm = as.numeric(pm)) %>%
  left_join(mc, by = "id") %>%
  filter(!is.na(gbm_pm)) %>%
  select(icsacach, ageyc, fmale, blck, whit, hisp, pcedu, 
         ipovl, pcemft, fthr, schbl, gbm_pm, intbl) %>%
  group_by(intbl) %>%
  arrange(gbm_pm) %>% # sort by group to ensure that the columns are matched
  nest() %>%
  pivot_wider(names_from = "intbl", values_from = "data", names_prefix = "g") %>%
  unnest(cols = c(g1, g0), names_sep = "_") %>%
  mutate(y = g1_icsacach - g0_icsacach,
         age = g1_ageyc - g0_ageyc,
         fmale = g1_fmale - g0_fmale,
         blck = g1_blck - g0_blck,
         whit = g1_whit - g0_whit,
         hisp = g1_hisp - g0_hisp,
         pcedu = g1_pcedu - g0_pcedu,
         ipovl = g1_ipovl - g0_ipovl,
         pcemft = g1_pcemft - g0_pcemft,
         fthr = g1_fthr - g0_fthr)
```

## Optimal pair matching Using pscores generated by Stata-boost (psb)

```{r error=F, warning=F, message=F}
prank <- rank(mc$psb)
names(prank) <- mc$id
d1 <- outer(prank[mc$intbl==1], prank[mc$intbl==0], "-")
d1 <- abs(d1)
pm1 <- pairmatch(d1, controls=1)
(pm1.d <- matched.distances(pm1,d1,pres=TRUE))
unlist(pm1.d,max)
mean(unlist(pm1.d))
sum(unlist(pm1.d))
stratumStructure(pm1)

# Merge to original dataset and calculate difference scores
m2.df <- tibble(id = as.numeric(names(pm1)), gbm_pm = as.numeric(pm1)) %>%
  left_join(mc, by = "id") %>%
  filter(!is.na(gbm_pm)) %>%
  select(icsacach, ageyc, fmale, blck, whit, hisp, pcedu, 
         ipovl, pcemft, fthr, schbl, gbm_pm, intbl) %>%
  group_by(intbl) %>%
  arrange(gbm_pm) %>% # sort by group to ensure that the columns are matched
  nest() %>%
  pivot_wider(names_from = "intbl", values_from = "data", names_prefix = "g") %>%
  unnest(cols = c(g1, g0), names_sep = "_") %>%
  mutate(y = g1_icsacach - g0_icsacach,
         age = g1_ageyc - g0_ageyc,
         fmale = g1_fmale - g0_fmale,
         blck = g1_blck - g0_blck,
         whit = g1_whit - g0_whit,
         hisp = g1_hisp - g0_hisp,
         pcedu = g1_pcedu - g0_pcedu,
         ipovl = g1_ipovl - g0_ipovl,
         pcemft = g1_pcemft - g0_pcemft,
         fthr = g1_fthr - g0_fthr)
```

## Balance Checks

```{r}
# Rand-gbm
bal.tab(pm, intbl ~ ageyc + fmale + blck + whit + hisp + 
          pcedu + ipovl + pcemft + fthr, data = mc, s.d.denom = "pooled")

# Stata-boost
bal.tab(pm1, intbl ~ ageyc + fmale + blck + whit + hisp + 
          pcedu + ipovl + pcemft + fthr, data = mc, s.d.denom = "pooled")
```

## Postmatching Regression Analysis Using Difference Scores

Results show that the treatment effects (i.e., constant of the regression model) in both models are not statistically significant. Thus, both methods provide similar findings.

```{r}
# Rand-gbm
m1 <- lm(y ~ age + fmale + blck + whit + hisp + pcedu + ipovl + 
                      pcemft + fthr, data = m1.df)
lmtest::coeftest(m1, vcov = vcovCL, cluster = m1.df$g0_schbl)

# Stata-boost
m2 <- lm(y ~ age + fmale + blck + whit + hisp + pcedu + ipovl + 
                      pcemft + fthr, data = m2.df)
lmtest::coeftest(m2, vcov = vcovCL, cluster = m2.df$g0_schbl)
```
