# Section 6.5.2 Subclassification & Cox Proportional Hazards (PS) 

```{r message=F, warning=F, error=F}
library(tidyverse)
library(haven)
library(MatchIt)
library(cobalt)
library(sjlabelled)
library(broom)
library(survival)
library(PSweight)
options(scipen=999)
```

## Import Data

```{r}
df <- read_dta("data/Data_Chapter6/sur_subclass.dta") %>%
  zap_formats()
df <- sjlabelled::remove_all_labels(df) %>%
  as_tibble()
```

## Create Propensity Score Using Logistic

```{r}
cov <- "married + high + bahigh + poverty2 + poverty3 + poverty4 + poverty5 + employ + open + black + hispanic + natam + chdage1 + chdage2 + chdage3 + cgrage1 + cgrage2 + cgrage3 + CRA47A + mental + arrest + PSH17A	+ sexual + provide + supervis + other + ra + cidi + cgneed"
f <- as.formula(paste("aodserv ~", cov))
m1 <- glm(aodserv ~ married + high + bahigh + poverty2 + poverty3 + 
          poverty4 + poverty5 + employ + open + black + hispanic + natam + chdage1 + 
          chdage2 + chdage3 + cgrage1 + cgrage2 + cgrage3 + CRA47A + mental + 
          arrest + PSH17A	+ sexual + provide + supervis + other + ra + cidi + cgneed,
          family = binomial,
          data = df)
ps3 <- predict(m1, type = "response")
sur_subclass1 <- df
sur_subclass1$ps3 <- ps3
```

## Function to Conduct Balance Check Using Normalized Difference

```{r}
nd <- function(cov, df = sur_subclass1, tx = "aodserv") {
  
  tx <- as.name(tx)
  cov <- as.name(cov)
  
  df <- df %>%
    group_by(!!tx) %>%
    summarize(m = mean(!!cov, na.rm = T),
              s = sd(!!cov, na.rm = T),
              .groups = "drop") %>%
    pivot_wider(names_from = !!tx, values_from = c("m", "s"), names_sep = "") %>%
    mutate(nd = (m1-m0)/sqrt(s1^2+s0^2))
  return(df$nd)
}
```

## Balance Check on the Original Sample (Normalized Difference)

```{r}
covs <- str_split(cov, "\\s[+]\\s")[[1]] # vector of covariates
sapply(covs, nd) # apply the function nd on each covariate
```

## Function to Check Balance Using Chi-Square Tests

```{r}
check_blc <- function(cov, df = sur_subclass1, tx = "aodserv") {
  ct <- tidy(
    chisq.test(
      table(df[[cov]], y = df[[tx]]), correct =F
      )
    ) %>%
    add_column(cov = cov, .before = "statistic") %>%
    mutate(p.value = round(p.value, 2))
  return(ct)
}
```

## Balance Check on the Original Sample (Chi-Square)

```{r warning=F}
covs2 <- c("married", "educ", "pov", "employ", "open", "race", "chdage",
           "cgage", "CRA47A", "mental", "arrest", "PSH17A", "maltx", "ra",
           "cidi", "cgneed")
map_dfr(covs2, check_blc)
```

## Get Quintiles

```{r}
q5 <- c(0, round(quantile(sur_subclass1$ps3, c(.2, .4, .6, .8), type = 1), 7), 1) # round to 7 digits to replicate Stata code
q5
```

## Create 5 Subclasses Using Quintiles

```{r}
# Cut 5 subclasses
subs <- sur_subclass1 %>%
  arrange(ps3) %>%
  group_by(q = cut(ps3, 
                   include.lowest = F,
                   right = F,
                   dig.lab = 7,
                   breaks = q5,
                   labels = F))

# Summary statistics of ps3 per subclass
subs %>%
  group_by(q) %>%
  summarise(obs = n(),
            mean = mean(ps3),
            sd = sd(ps3),
            min = min(ps3),
            max = max(ps3), .groups = "drop")

# Filter each subclass into its own tibble
sub1 <- subs %>% filter(q == 1)
sub2 <- subs %>% filter(q == 2)
sub3 <- subs %>% filter(q == 3)
sub4 <- subs %>% filter(q == 4)
sub5 <- subs %>% filter(q == 5)
```

## Check Imbalance Post-Subclassification (5 Subclasses)

```{r warning=F}
map_dfr(covs2, check_blc, df = sub1)
map_dfr(covs2, check_blc, df = sub2)
map_dfr(covs2, check_blc, df = sub3)
map_dfr(covs2, check_blc, df = sub4)
map_dfr(covs2, check_blc, df = sub5)
```

## Trimming: Search for Best Cutoff Alpha for Overlap

```{r}
# Program to search best cutoff value alpha for overlap (Crum et al., 2009)
trim <- PSweight::PStrim(
  data = as.data.frame(sur_subclass1), # fix for latest package update
  ps.formula = f,
  optimal = T
)
trim$delta # or use .079 from the Stata results
1-trim$delta # or use .921 from the Stata results
```

## Trim Observations

```{r}
summary(sur_subclass1$ps3)
sur_subclass2 <- sur_subclass1 %>%
  filter(ps3 >= .079 & ps3 <= .921)
summary(sur_subclass2$ps3)
```

## Create 5 New Subclasses and Check Imbalance

```{r warning=F}
# Cut 5 subclasses (Type 2 and 5 in Quantile() are identical to Stata results)
subs_t <- sur_subclass2 %>%
  arrange(ps3) %>%
  group_by(q = cut(ps3, 
                   include.lowest = F,
                   right = F,
                   dig.lab = 7,
                   breaks = c(0, round(quantile(sur_subclass2$ps3, 
                            c(.2, .4, .6, .8), type = 2), 7), 2),
                   labels = F))

# Summary statistics of ps3 per subclass
subs_t %>%
  group_by(q) %>%
  summarise(obs = n(),
            mean = mean(ps3),
            sd = sd(ps3),
            min = min(ps3),
            max = max(ps3), .groups = "drop")

# Filter each subclass into its own tibble
sub1_t <- subs_t %>% filter(q == 1)
sub2_t <- subs_t %>% filter(q == 2)
sub3_t <- subs_t %>% filter(q == 3)
sub4_t <- subs_t %>% filter(q == 4)
sub5_t <- subs_t %>% filter(q == 5)

# Check imbalance
map_dfr(covs2, check_blc, df = sub1_t)
map_dfr(covs2, check_blc, df = sub2_t)
map_dfr(covs2, check_blc, df = sub3_t)
map_dfr(covs2, check_blc, df = sub4_t)
map_dfr(covs2, check_blc, df = sub5_t)
```

## Outcome Analysis by Subclass

```{r}
f2 <- as.formula("Surv(durm, rrpt) ~ aodserv + married + high + bahigh + poverty2 + poverty3 + poverty4 + poverty5 + employ + open + black + hispanic + natam + chdage1 + chdage2 + chdage3 + CRA47A")
tidy(survival::coxph(f2, sub1_t))
tidy(survival::coxph(f2, sub2_t))
tidy(survival::coxph(f2, sub3_t))
tidy(survival::coxph(f2, sub4_t))
tidy(survival::coxph(f2, sub5_t))
```

## Aggregate Results

```{r}
exp(0.37513328)
(1-pnorm(1.899353989))*2 # p = .05751795
tidy(survival::coxph(f2, df))
```

## Subclassification with MatchIt

```{r}
mm <- MatchIt::matchit(f, df, method = "subclass", subclass = 5)
love.plot(mm, binary = "std", disp.subclass = T)
bal.tab(mm, un = T, binary = "std", disp.subclass = T)
```
