# Section 7.3.1 Propensity Score Weighting (SD & PS)

## Preparation

```{r message=F, warning=F, error=F}
library(WeightIt)
library(broom)
library(cobalt)
library(sandwich)
library(stargazer)
library(tidyverse)
library(lmtest)

# Load data
d <- haven::read_dta("data/Data_Chapter7/chpt5_2_original.dta")
```

## Solution 1: Obtain Weights with WeightIt

### Weights 

```{r message=FALSE, warning=FALSE}
# ATE weight
w1.out <- WeightIt::weightit(kuse ~ male + black + age97 + pcged97 + mratio96,
                             data = d,
                             ps = d$ps,
                             estimand = "ATE")
summary(w1.out)

# ATT weight
w2.out <- WeightIt::weightit(kuse ~ male + black + age97 + pcged97 + mratio96,
                             data = d,
                             ps = d$ps,
                             estimand = "ATT")
summary(w2.out)
```

Visual balance checks may be carried out with:

```{r eval=F}
cobalt::bal.tab(w1.out, m.threshold = .05, disp.v.ratio = TRUE)
cobalt::bal.tab(w2.out, m.threshold = .05, disp.v.ratio = TRUE)
```

### PSW With ATE Weights

```{r message=FALSE, warning=FALSE}
m1 <- lm(lwss97~ kuse + male + black + age97 + pcged97 + mratio96, data = d, weights = w1.out$weights) 
lmtest::coeftest(m1, vcov. = vcovCL(m1, cluster = d$pcg_id))
```

### PSW With ATT Weights

```{r message=FALSE, warning=FALSE}
m2 <- lm(lwss97~ kuse + male + black + age97 + pcged97 + mratio96, data = d, weights = w2.out$weights)
lmtest::coeftest(m2, vcov. = vcovCL(m2, cluster = d$pcg_id))
```

## Solution 2: Creating the Weights Manually 

### Weights

```{r message=FALSE, warning=FALSE}
d.weights <- d %>%
  mutate(ate_w = ifelse(kuse == 0, 1/(1-ps), 1/ps),
         att_w = ifelse(kuse == 0, ps/(1-ps), 1))
```

### PSW With ATE Weights

```{r message=FALSE, warning=FALSE}
# psw with ate_w weights
m3 <- lm(lwss97~ kuse + male + black + age97 + pcged97 + mratio96, data = d.weights, weights = ate_w) 
lmtest::coeftest(m3, vcov. = vcovCL(m3, cluster = d.weights$pcg_id))
```

### PSW With ATT Weights

```{r message=FALSE, warning=FALSE}
m4 <- lm(lwss97~ kuse + male + black + age97 + pcged97 + mratio96, data = d.weights, weights = att_w)
lmtest::coeftest(m4, vcov. = vcovCL(m4, cluster = d.weights$pcg_id))
```

## Imbalance Checks

The balance check used regression for a continuous dependent variable and logistic regression for a dichotomous dependent variable.

```{r}
c1 <- glm(male ~ kuse, family = quasibinomial, data = d.weights, weights = ate_w)
lmtest::coeftest(c1, vcov. = vcovCL(c1, cluster = d.weights$pcg_id))
c2 <- glm(male ~ kuse, family = quasibinomial, data = d.weights, weights = att_w)
lmtest::coeftest(c2, vcov. = vcovCL(c2, cluster = d.weights$pcg_id))
c3 <- glm(black ~ kuse, family = quasibinomial, data = d.weights, weights = ate_w)
lmtest::coeftest(c3, vcov. = vcovCL(c3, cluster = d.weights$pcg_id))
c4 <- glm(black ~ kuse, family = quasibinomial, data = d.weights, weights = att_w)
lmtest::coeftest(c4, vcov. = vcovCL(c4, cluster = d.weights$pcg_id))

c5 <- lm(age97 ~ kuse, weights = ate_w, data = d.weights)
lmtest::coeftest(c5, vcov. = vcovCL(c5, cluster = d.weights$pcg_id))
c6 <- lm(age97 ~ kuse, weights = att_w, data = d.weights)
lmtest::coeftest(c6, vcov. = vcovCL(c6, cluster = d.weights$pcg_id))
c7 <- lm(pcged97 ~ kuse, weights = ate_w, data = d.weights)
lmtest::coeftest(c7, vcov. = vcovCL(c7, cluster = d.weights$pcg_id))
c8 <- lm(pcged97 ~ kuse, weights = att_w, data = d.weights)
lmtest::coeftest(c8, vcov. = vcovCL(c8, cluster = d.weights$pcg_id))
c9 <- lm(mratio96 ~ kuse, weights = ate_w, data = d.weights)
lmtest::coeftest(c9, vcov. = vcovCL(c9, cluster = d.weights$pcg_id))
c10 <- lm(mratio96 ~ kuse, weights = att_w, data = d.weights)
lmtest::coeftest(c10, vcov. = vcovCL(c10, cluster = d.weights$pcg_id))
```
