# Section 8.4.1 Matching with Bias-Corrected and Robust Variance Estimators (SD & PS)

## Preparation

```{r message=F, warning=F, error=F}
library(lmtest)
library(tidyverse)
library(broom)
library(Matching)

# Load Data
d <- haven::read_dta("data/Data_Chapter8/cds_pcss97.dta")
```

## Breusch-Pagan Test for Heteroskedasticity

```{r message=FALSE, warning=FALSE}
m0 <- lm(pcss97 ~ kuse + male + black + age97 + pcged97 + mratio96 + pcg_adc, data = d)
get_bptest <- function(data, lm.model, var) {
  b <- lmtest::bptest(lm.model, as.formula(paste0("~", var)), 
                      data = data, studentize = F)
  return(tibble(variable = var, statistic = b$statistic, 
                df = b$parameter, p.value = b$p.value))
}
map_dfr(c("kuse", "male", "black", "age97", "pcged97", "mratio96", "pcg_adc"), 
        get_bptest, data = d, lm.model = m0)
```

## Matching Estimators

### Sample Average Treatment Effect (SATE)

```{r message=FALSE, warning=FALSE}
match_sate <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATE",
  M = 4,
  BiasAdjust = TRUE,
  sample = TRUE,
  Var.calc = 4
  ))
summary(match_sate)
```

### Population Average Treatment Effect (PATE)

```{r message=FALSE, warning=FALSE}
match_pate <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATE",
  M = 4,
  BiasAdjust = TRUE,
  sample = FALSE,
  Var.calc = 4
  ))
summary(match_pate)
```

### Sample average treatment effect for the treated (SATT)

```{r message=FALSE, warning=FALSE}
match_satt <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATT",
  M = 4,
  BiasAdjust = TRUE,
  sample = TRUE,
  Var.calc = 4
))
summary(match_satt)
```

### Population average treatment effect for the treated (PATT)

```{r message=FALSE, warning=FALSE}
match_patt <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATT",
  M = 4,
  BiasAdjust = TRUE,
  sample = FALSE,
  Var.calc = 4
))
summary(match_patt)
```

### Sample average treatment effect for the controls (SATC)

```{r message=FALSE, warning=FALSE}
match_satc <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATC",
  M = 4,
  BiasAdjust = TRUE,
  sample = TRUE,
  Var.calc = 4
))
summary(match_satc)
```

### Population average treatment effect for the controls (PATC)

```{r message=FALSE, warning=FALSE}
match_patc <- with(d, Matching::Match(
  Y = pcss97,
  Tr = kuse,
  X = data.frame(male, black, age97, pcged97, mratio96, pcg_adc),
  estimand = "ATC",
  M = 4,
  BiasAdjust = TRUE,
  sample = FALSE,
  Var.calc = 4
))
summary(match_patc)
```
